Symbolic Abstract Data Type Inference

### 2015-07-07 motivation.tex
Presentation of modular reasoning.

Effective scalable reasoning about nontrivial software implementations generally
requires considering each software module separately, in isolation, using
abstract specifications for other modules. When modules are objects whose
methods may be called concurrently, their behavior is typically understood in
terms of valid invocation sequences of atomic abstract data types (ADTs): any
execution of overlapping method invocations can be \emph{linearized} into an
invocation sequence admitted by the ADT~\cite{journals/toplas/HerlihyW90}. For
example, consider the execution history depicted in Figure~\ref{fig:clients} in
which the add operations numbered $2$ and $3$ overlap with each other, and,
respectively, with operations $1$ and $4$. Among the five possible ways to
linearize these six operations, the linearization~$1$; $3$; $2$; $4$; $5$; $6$
does correspond to a valid sequence of the atomic queue ADT, and so this history
is linearizable. ADT specifications thus decouple reasoning about object
implementations — is each concurrent execution linearizable? — from their
clients’ invocations — do all valid linearizations preserve client invariants?

>>>>

Effective scalable reasoning about nontrivial software implementations generally
requires considering each software module separately, in isolation, using
abstract specifications for other modules. When modules are objects whose
methods may be called concurrently, their behavior is typically understood in
terms of valid invocation sequences of atomic abstract data types (ADTs): any
execution of overlapping method invocations can be \emph{linearized} into an
invocation sequence admitted by the ADT~\cite{journals/toplas/HerlihyW90}. For
example, consider the execution history depicted in Figure~\ref{fig:clients} in
which the add operations numbered $2$ and $3$ overlap with each other, and,
respectively, with operations $1$ and $4$. Among the five possible ways to
linearize these six operations, the linearization~$1$; $3$; $2$; $4$; $5$; $6$
does correspond to a valid sequence of the atomic queue ADT, and so this history
is linearizable. ADT specifications thus decouple reasoning about object
implementations from their client’s invocations:
\begin{itemize}

  \item Does every implementation execution have a valid linearization?

  \item Does every valid linearizations preserve client invariants?

\end{itemize}
The former question depends only on a given object’s implementation, and the
latter only on a given object’s clients.
